\section{Experience}
\cventry{Aug. 2019-}
{Software developer apprentice}
{Waystar}
{Louisville, KY}
{}
{Technologies used daily include: 
	C\# with Visual Studio integrated development environment (IDE); 
	Entity Framework, LINQ, and Microsoft SQL Server Management Studio (SSMS); 
	Git with Azure DevOps/TFS;  
	JavaScript with jQuery and Ajax, 
	and HTML5 and CSS3 with Bootstrap. 
	Currently contributing to a company-wide .NET framework applicant testing website implementing MVC pattern and Onion Architecture.}
\cventry{2017-2018}
{Ontology, Latin, and classical political philosophy teacher}
{Highlands Latin School}
{Louisville, KY}
{}
{Taught college-level material in Latin, ontology, and classics at advanced private high school. 
	Led ontology students to fully reconstruct arguments 
	and critically explicate texts from Plato, Aristotle, Thomas Aquinas, and others. 
	Led Latin students to translate both from and to Latin, 
	with assignments principally focused on Caesar. 
	By pairing classic texts with carefully designed assignments, 
	I improved student learning outcomes 
	while decreasing student workload, 
	leading 90\% of students to earn the highest letter grade available.}
\cventry{2015-2017}
{Distinguished research fellow}
{Fordham University}
{Bronx, NY}
{}
{Published numerous \href{https://fordham.academia.edu/jacobarchambault}{articles}, \href{https://www.collegepublications.co.uk/dialogues/?00005}{book} \href{https://www.cambridgescholars.com/hylomorphism-and-mereology}{contributions}, and an \href{https://brill.com/view/journals/viv/56/3-4/viv.56.issue-3-4.xml?language=en}{edited volume }
	with publishers including Cambridge University Press, Brill and Springer, 
	and presented papers at prestigious institutions across three continents, 
	primarily in the history and philosophy of logic. 
	For a full list of publications and presentations, see my 
	\href{https://jacobarchambault.com/curriculum-vitae}{\emph{CV}}.}
