\subsection{Venminder}
%\datedsubsectionnarrow{Feb 2023-present}{Production support team}{Software Developer II}{
%}
\datedsubsectionnarrow{May 2022-Feb 2023}{Control assessments team}{Software Developer II}{ 
	Created Typescript Aurelia components, 
	C\# immutable records, 
	sealed request and command handlers, 
	Xunit tests, 
	and DynamoDB tables via AWS CloudFormation 
	for control assessments containerized microservice instantiating CQRS architecture, 
	and added generic batch update ability to our Dynamo data access library using AWS' SDK for .NET.
}
\subsection{PetPartners}
\datedsubsectionnarrow{Oct 2021-May 2022}{Production support team}{Associate Software Developer}{
	Consolidated repeat API calls and removed Cartesian explosion for premium tracking feature causing slow coverage period loading, leading to a 26X performance improvement. 
	Wrote wiki documentation for the company's release process, 
	git workflow, 
	and numerous routine bugfixes,  
	and created Azure DevOps bug template subsequently adopted across entire company.
}
\datedsubsectionnarrow{April-Oct 2021}{Backend team}{}{
	Created backend logic, 
	C\# web API endpoints, 
	and normalized SQL tables via EF migrations for handling commission payments to subcontracted companies and internal call center agents for new flagship employer-based pet insurance product. 
	Converted SQL stored procedures for retrieving claims, 
	policy losses, 
	and veterinarian information in monolithic application 
	to Entity Framework backend called by web API microservices,  
	and scripted automation of .NET web API publication for local development environments. 
}
\subsection{Waystar}
\datedsubsectionnarrow{Jan-April 2021}{Recondo Bots team}{Software Developer Apprentice}
{Fixed bugs in Chrome and Firefox web crawlers, written as Java Maven projects using Selenium, obtaining patient payment information from insurance provider websites. Deployed fixes to development and production environments with Jenkins. Used Jira for project management, SVN for version control, and Eclipse and Intellij Idea for IDEs.
}
\datedsubsectionnarrow{May 2020-April 2021}{DevOps Applications team}{}
{Louisville, KY. Pioneered work converting .NET Framework projects to containerized .NET Core projects running on Docker; built, deployed, and configured team projects with Azure DevOps; set up logging of automated alerts from A10 load balancers to Microsoft Teams and Splunk, and co-led company-wide training on using YAML for Azure DevOps pipeline configuration.
}
\datedsubsectionnarrow{Aug 2019-May 2020}
{Architecture team}
{}
{Louisville, KY. Contributed to a company-wide full-stack C\# .NET framework applicant testing website implementing MVC pattern with Visual Studio and Git version control. Normalized back-end SQL Server databases and application model layer, and wrote database queries both in SQL directly and in C\# with LINQ and Entity Framework. Wrote JavaScript/jQuery functions for DOM manipulation and sending/retrieving server information with AJAX, and improved HTML/CSS web accessibility and design responsiveness by adding semantic HTML, media queries, and Flexbox layouts. }
