\section{Employment}
\cventry{May 2020-}{Software developer apprentice, DevOps Applications team}{Waystar}{Louisville, KY}{}
{Led work on porting various team projects from .NET Framework to .NET Core and on project containerization with Docker; built, deployed, and configured team projects with Azure DevOps; set up logging of automated alerts from A10 load balancers to Microsoft Teams and Splunk, and co-led company-wide training on using YAML for Azure DevOps pipeline configuration.
}
\cventry{Aug. 2019-May 2020}
{Software developer apprentice, Architecture team}
{}
{}
{}
{Contributed to a company-wide full-stack C\# .NET framework applicant testing website implementing MVC pattern with Visual Studio and Git version control. Normalized back-end SQL Server databases and application model layer, and wrote database queries both in SQL directly and in C\# with LINQ and Entity Framework. Wrote JavaScript/jQuery functions for DOM manipulation and sending/retrieving server information with AJAX, and improved HTML/CSS web accessibility and design responsiveness by adding semantic HTML, media queries, and Flexbox layouts. 
	}
\cventry{2017-2018}
{Ontology, Latin, and classical political philosophy teacher}
{Highlands Latin School}
{Louisville, KY}
{}
{Taught college-level material in Latin, ontology, and classics at advanced private high school. 
	Led ontology students to fully reconstruct arguments 
	and critically explicate texts from Plato, Aristotle, Thomas Aquinas, and others. 
	Led Latin students to translate both from and to Latin, 
	with assignments principally focused on Caesar. 
	By pairing classic texts with carefully designed assignments, 
	I improved student learning outcomes 
	while decreasing student workload, 
	leading 90\% of students to earn the highest letter grade available.}
\cventry{2015-2017}
{Distinguished research fellow}
{Fordham University}
{Bronx, NY}
{}
{Published numerous \href{https://fordham.academia.edu/jacobarchambault}{articles}, \href{https://www.collegepublications.co.uk/dialogues/?00005}{book} \href{https://www.cambridgescholars.com/hylomorphism-and-mereology}{contributions}, and an \href{https://brill.com/view/journals/viv/56/3-4/viv.56.issue-3-4.xml?language=en}{edited volume }
	with publishers including Cambridge University Press, Brill and Springer, 
	and presented papers at prestigious institutions across three continents, 
	primarily in the history and philosophy of logic. 
	For a full list of publications and presentations, see my 
	\href{https://jacobarchambault.com/curriculum-vitae}{\emph{CV}}.}
