\subsection{Conferences}
\begin{enumerate}
	\setcounter{enumi}{\value{presentationCounter}}	% Continue counting
\item \datedsubsectionnarrow{2019}{}{`Dialectically situating medieval accounts of the subject matter of logic'. International Medieval Congress. University of Leeds. Leeds, UK. July 1-4.}{}

\item \datedsubsectionnarrow{2018}{}{`The semantic account of formal consequence, from Alfred Tarski back to John Buridan'. Indiana Philosophical Association, Fall 2018 meeting. Indiana University Bloomington. Bloomington, IN. November 16-17.}{} 
\item \datedsubsectionnarrow{}{}{`Divided modality for Ockhamists'. 6th World Congress on Universal Logic. Campus Albert Londres. Vichy, France, June 21-26.}{} 
\item \datedsubsectionnarrow{2017}{}{`Counterpossible conditionals in the Filioque Controversy'. 2nd World Congress on Logic and Religion. University of Warsaw. Warsaw, Poland. June 18-22.}{} 
\item \datedsubsectionnarrow{2016}{}{`Consequence and formality in the logic of Walter Burley'. HOPOS 2016: Eleventh International Congress of the International Society for the History of Philosophy of Science. University of Minnesota Twin Cities. Minneapolis-St. Paul, MN. June 22-25.}{} 
\item \datedsubsectionnarrow{}{}{`Mereological hylomorphism and the development of the Buridanian account of formal consequence'. International Congress on Medieval Studies. Western Michigan University. Kalamazoo, MI. May 12-15.}{} 
\item \datedsubsectionnarrow{}{}{`Deflationism, liars, and the ontic use of `true''. Long Island Philosophical Society Spring 2016 Conference. Molloy College. Rockville Centre, NY. April 9.}{} 
\item \datedsubsectionnarrow{}{}{`Distinguishing Leibniz's system of pre-established harmony from the system of occasional causes'. The Continuing Relevance of Leibniz. Franciscan University of Steubenville. Steubenville, OH. April 8-9.}{} 
\item \datedsubsectionnarrow{2015}{}{`Model-theoretic inferentialism for bilateral natural deduction'. Inferentialism Workshop. University of St. Andrews. St. Andrews, United Kingdom. November 25-26.}{} 
\item \datedsubsectionnarrow{}{}{`Three medieval accounts of the subject matter of logic'. 33rd Annual Meeting of the Society for Ancient Greek Philosophy with the Society for the Study of Islamic Philosophy and Science. Fordham University. New York, NY. October 23-25.}{} 
\item \datedsubsectionnarrow{}{}{`Interpretations of quantifiers and logical hylomorphism'. Entia et Nomina V Workshop. Jagiellonian University. Krakow, Poland. September 9-11.}{} 
\item \datedsubsectionnarrow{}{}{`The teaching of the trivium at Bec and its bearing upon the Anselmian program of fides quaerens intellectum'. Reading Anselm: Context and Criticism. Boston College. Boston, MA. July 27-30.}{} 
\item \datedsubsectionnarrow{}{}{`The monadothergy'. Emmanuel Levinas across the Generations and Continents: North American Levinas Society Annual Meeting 2015. Purdue University. Lafayette, IN. July 27-30.}{} 
\item \datedsubsectionnarrow{}{}{`The teaching of the trivium at Bec and its bearing upon the Anselmian programme of fides quaerens intellectum'. International Medieval Congress. University of Leeds. Leeds, United Kingdom. July 6-9.}{} 
\item \datedsubsectionnarrow{}{}{`John Buridan on forms and formal logic'. Durham Day of Medieval Thought. Durham University. Durham, United Kingdom. July 5.}{} 
\item \datedsubsectionnarrow{}{}{`Interpretations of quantifiers and logical hylomorphism'. Aristotelian Solutions to Contemporary Problems. Indiana University-Purdue University, Fort Wayne. Fort Wayne, IN. April 25-26.}{} 
\item \datedsubsectionnarrow{}{}{`Is the reticulational model of scientific change based on a misunderstanding? Revisiting the Laudan-Kuhn debate'. Long Island Philosophical Society Conference. St. John's University. Queens, NY. April 18.}{} 
\item \datedsubsectionnarrow{}{}{`Monotonic and non-monotonic variations on Anselm's proof'. 1st World Congress on Logic and Religion. Hotel Tropical Tambaú. João Pessoa, Paraíba, Brazil. April 1-5.}{} 
\item \datedsubsectionnarrow{2014}{}{`Consequence and proportionality in Boethius' De Topicis Differentiis'. Analogical Reasoning – East and West: Formal Models and Practical Applications. Universität Heidelberg. Heidelberg, Germany. November 24-25.}{} 
\item \datedsubsectionnarrow{}{}{`Leibnizian intelligibility'. Eighth annual conference of the Leibniz Society of North America. University of South Florida. Tampa, Florida. October 31-November 2.}{} 
\item \datedsubsectionnarrow{}{}{`Authority, order of reading, and authenticity in two medieval manuscripts of the Logica Vetus'. International Medieval Congress. University of Leeds. Leeds, United Kingdom. July 7-10.}{} 
\item \datedsubsectionnarrow{2013}{}{`On Leibnizian harmony and Malebranchian occasionalism: Leibniz's Cartesian route out of Cartesian dualism'. Theistic Metaphysics and Naturalism: Society of Christian Philosophers 2013 Eastern Division Conference. University of South Florida. Tampa, FL. October 24-26.}{} 
\item \datedsubsectionnarrow{2012}{}{`Nature, will, and the Fall in Augustine and Maximus the Confessor'. Divine and Human Action: Agency and Action in Philosophy and Theology. Houston Baptist University. Houston, TX. March 30-31.}{} 
\item \datedsubsectionnarrow{2009}{}{`Agnosia and henosis in Pseudo-Dionysius the Areopagite'. Neoplatonism and its Legacy. Franciscan University. Steubenville, OH. April 25.}{} 
	\setcounter{presentationCounter}{\value{enumi}}	% Remember the counts
\end{enumerate}
\subsection{Workshops and discussion groups}
\begin{enumerate}
	\setcounter{enumi}{\value{presentationCounter}}	% Continue counting
	\item \datedsubsectionnarrow{2016}{}{`The semantic account of formal consequence, from Alfred Tarski back to John Buridan'. Logic and Metaphysics Workshop. CUNY Graduate Center. New York, NY. Oct 31.}{} 
	\item \datedsubsectionnarrow{2014}{}{`Minimalism, liars, and pre-sentential truth'. Models, Modality, and Meaning Group. University of St. Andrews. St. Andrews, United Kingdom. November 6.
	}{} 
	\setcounter{presentationCounter}{\value{enumi}}	% Remember the counts
\end{enumerate}
\subsection{Discussant}
\begin{enumerate}
	\setcounter{enumi}{\value{presentationCounter}}	% Continue counting
	\item \datedsubsectionnarrow{2018}{}{Seungil Lee, `The rate of time's passage and special relativity'. Indiana Philosophical Association, Fall 2018 meeting. Indiana University Bloomington. Bloomington, IN. November 16-17.}{} 
	\item \datedsubsectionnarrow{2015}{}{Joanna Luc, `Necessary and accidental qualities in CIFOL'. Entia et Nomina V Workshop. Jagiellonian University. Krakow, Poland. September 9-11.}{} 
	\setcounter{presentationCounter}{\value{enumi}}	% Remember the counts	
\end{enumerate}
\subsection{Graduate Conferences}
\begin{enumerate}
\setcounter{enumi}{\value{presentationCounter}}	% Continue counting
	\item \datedsubsectionnarrow{2016}{}{`Powers presentism'. The Philosophy of Time: 2016 Graduate Gateway Conference. University of Missouri, St. Louis. St. Louis, MO. April 23-24.}{} 
	\item \datedsubsectionnarrow{2014}{}{`The monadothergy'. Philosophy on Trial: The Others of Philosophy. DePaul University Graduate Philosophy Conference. DePaul University. Chicago, IL. February 14-15.}{} 
\setcounter{presentationCounter}{\value{enumi}}	% Remember the counts	
\end{enumerate}
