\section{Service}
\subsection{Manuscript work}
Identification of the text of \href{https://digital-scriptorium.org/xtf3/search?rmode=digscript;smode=basic;text=Jacob\%20Archambault;docsPerPage=1;startDoc=1;fullview=yes}{Smith Medieval/Renaissance Fragment 27}. Columbia University, Rare Book and Manuscript Library. .
\subsection{Referee}
\begin{multicols}{2}
- Australasian Journal of Philosophy

- Journal of Pure and Applied Logic

- Kriterion

\columnbreak
- Logica Universalis\\
- Vivarium
\end{multicols}

\subsection{Session chair}
\begin{enumerate}
	\item \datedsubsectionnarrow{2017}
{}
{Logic and Western Christianity. 2nd World Congress on Logic and Religion. University of Warsaw. Warsaw, Poland. June 18-22.}
{}
\item \datedsubsectionnarrow{2016}
{}
{Symposium: Post-Aristotelian Science in the Fourteenth Century. HOPOS 2016: Eleventh International Congress of the International Society for the History of Philosophy of Science. University of Minnesota Twin Cities. Minneapolis-St. Paul, MN. June 22-25.}
{}
\item \datedsubsectionnarrow{}
{}
{Hans Thomas Adriaenssen, ‘Substantial forms between physics and metaphysics. John Sergeant and Antoine Le Grand’; and Henrik Lagerlund, ‘The mechanization of philosophy between 1300-1700: some reflections’. Workshop on Medieval Natural Philosophy. University of St. Thomas. St. Paul, MN. June 22.}
{}
\item \datedsubsectionnarrow{2015}
{}
{Thomas Kivatinos. ‘Humean metaphysics and grounding: a peculiar alliance’. The Metaphysics and Epistemology of Grounding and Fundamentality. CUNY Graduate Center. New York, NY. December 10-11.}
{}
\item \datedsubsectionnarrow{}
{}
{Maya Eddom. ‘Intrinsic explanations and numerical representations’. The Metaphysics of Quantity. New York University. New York, NY. May 1-3, 2015.}
{}
\end{enumerate}
\subsection{Research assistant}
\datedsubsectionnarrow{2011-2012}
{Fordham University}
{Research Assistant to Giorgio Pini and Christopher Cullen.}
{}
\subsection{Professional affiliations}
- American Philosophical Association

- International Society for the History of Philosophy of Science
	
- Society for Ancient Greek and Islamic Philosophy

- Society for Medieval Logic and Metaphysics	

